---
editor_options:
  chunk_output_type: console
output:
  word_document: default
  html_document: default
---
# Module 5 - Assignment 1
## Reyes, Marielli Nicole
### Data Transformation

```{r}
library(tidyverse)
state_income <- read_csv("state_income.csv")
```

#### State Incomes
In this section of the project, a subset of data from the state income data file will be created. The following variables will be included: (1) State_Name, a character data type, which states the name of the state of the identified geographic location reported by the U.S. Census Bureau, (2) State_ab, a character data type, that includes the abbreviated name of the state reported by the U.S. Census Bureau for the identified geographic location, (3) County, a character data type, that states the name of the county reported by the U.S. Census Bureau for the identified geographic location, (4) City, a character data type, that includes the name of the city reported by the U.S. Census Bureau for the identified geographic location, (5) Type, a character data type, that states the type of place reported by the U.S. Census Bureau for the identified geographic location, (6) Aland, a double data type, that includes the square area of land at the identified geographic location, (7) Mean, a double data type, that states the mean household income for the identified geographic location, (8) Median, a double data type, that includes the median household income of the identified geographic location, and (9) Stdev, a double data type, that states the standard deviation of the household income of the identified geographic location. 

```{r}
state_income2 <- select(state_income,State_Name,State_ab,County,City,Type,ALand,Mean,Median,Stdev)

state_income2 <- select(state_income2,State_ab,everything())
head(state_income2)
```

```{r}
state_income2 <- rename(state_income2,SquareArea=ALand,IncomeMean=Mean,IncomeMedian=Median,IncomeStDev=Stdev)
head(state_income2)
```

```{r}
NC_income <- filter(state_income2,State_ab=="NC")
head(NC_income)
```


#### NC Incomes
In this particular section, the NC_income dataset will be used to come up with summaries of the incomes within North Carolina, which include summaries by county, city and type.

```{r}
NC_income <- arrange(NC_income,County)
head(NC_income)
```


```{r}
summary1 <- group_by(NC_income,County)
summary1 <- summarise (summary1, mean=mean(IncomeMean))

summary2 <- NC_income %>%
 group_by(City) %>%
 summarise (mean = mean(IncomeMean))
```

The summary1 dataset is comprised of the mean household incomes of the counties in NC, while the summary2 dataset show the mean household incomes of the cities in NC.

```{r}
summary3 <- NC_income %>%
  group_by(Type) %>%
  summarise (mean=mean(IncomeMean))
```

The summary3 dataset show the mean household incomes of the type of properties in NC.

#### Income Visualization
```{r}
ggplot(summary1, aes(x=County,y=mean))+
    geom_point()+
    theme(axis.text.x = element_text(angle = 90,vjust =0.5, hjust=1))+
    labs(title="County Income",
             y="Income") 
```

Based from the data, the county with the highest average income is Wake County and the county with the lowest average income is Stokes County. Moreover, the county with the second lowest average income is Surry County. 

```{r}
AvgStateIncome <- state_income2 %>%
  group_by(State_ab) %>%
  summarise(mean=mean(IncomeMean))

AvgStateIncome <- filter(AvgStateIncome,grepl("N",State_ab))

ggplot(AvgStateIncome, aes(x=State_ab,y=mean))+
    geom_point()+
    labs(title="State Income",
             x="State",
             y="Income") 
```

The scatterplot shows that New Jersey had the largest average income while Tennessee had the smallest average income. 
